A Plasma Torus Around a Young Low-Mass Star

### 2025-03-14 paper/ms.tex
update

{\it Effective temperature, radius, mass, and spectral type} We adopted
the star's effective temperature and radius measured using the spectral
energy distribution (SED) fitting procedure described by
\cite{Bouma2024}.  To summarize, this approach used
\texttt{astroARIADNE} \cite{Vines2022} with the BT-Settl stellar
atmosphere models \cite{Allard2012}, assuming the \cite{Asplund2009}
solar abundances and the \cite{Barber2006} water line lists.  This
approach fitted for the stellar effective temperature, radius,
reddening, surface gravity, metallicity, and distance by comparing the
measured broadband magnitudes against pre-computed model grids.
Specifically, we used the broadband magnitudes from Gaia DR2, APASS,
2MASS, SDSS, and WISE $W1$ and $W2$.  The resulting best-fit SED is
shown in Figure~\ref{fig:sed}.  This method has the most constraining
power for the star's effective temperature (2972 $\pm$ 40\,K) and radius
(0.42 $\pm$ 0.02\,$R_\odot$); the surface gravity, metallicity, and
reddening are typically only weakly constrained.  We determined the
star's spectral type to be M5.5Ve by visually comparing against the
photometric standards tabulated by \cite{Bochanski2007}, which agrees
with the effective temperature from the SED fitting \cite{Pecaut2013}.

@@

{\it Upper and Low Bounds on Dust}---An upper limit on the amount of hot
dust follows from the lack of an infrared excess.  A lower limit follows
from the assumption that most of the broadband optical depth comes from 
dust, rather than some other source.
Figure~\ref{fig:sed} shows the SED; ...
%TODO TODO LEFT OFF HERE
...

@@

\item[Acknowledgments]
  % convos w/
  % M.~Jardine. (line core sum, explaining flux tubes)
  % A Weinberger (dissolving rocks)
  % L Hillenbrand (general?)
  % line profile analysis by B.~Tofflemire.
  % CPS group for reduction support.
  The author thanks X, Y, Z.
  L.G.B. was suported by...
	Acknowledge TESS...

>>>>

{\it Effective temperature, radius, mass, and spectral type} We adopted
the star's effective temperature and radius measured using the spectral
energy distribution (SED) fitting procedure described by
\cite{Bouma2024}.  To summarize, this approach used
\texttt{astroARIADNE} \cite{Vines2022} with the BT-Settl stellar
atmosphere models \cite{Allard2012}, assuming the \cite{Asplund2009}
solar abundances and the \cite{Barber2006} water line lists.  This
approach fitted for the stellar effective temperature, radius,
reddening, surface gravity, metallicity, and distance by comparing the
measured broadband magnitudes against pre-computed model grids.
Specifically, we performed the fit using the broadband magnitudes from
Gaia DR2, APASS, 2MASS, SDSS, and WISE $W1$ and $W2$.  The resulting
best-fit SED is shown in Figure~\ref{fig:sed}.  This method has the most
constraining power for the star's effective temperature (2972 $\pm$
40\,K) and radius (0.42 $\pm$ 0.02\,$R_\odot$); the surface gravity,
metallicity, and reddening are typically only weakly constrained.  We
determined the star's spectral type to be M5.5Ve by visually comparing
the HIRES spectra against the photometric standards tabulated by
\cite{Bochanski2007}, which agrees with the effective temperature from
the SED fitting \cite{Pecaut2013}.

@@

{\it Upper and Low Bounds on Dust}---An upper limit on the amount of hot
dust follows from the lack of an infrared excess.  A lower limit follows
if one assumes that most of the broadband optical depth comes from
absorption and scattering by dust, rather than any radiative processes
associated with the plasma.

How much dust is allowed, given the lack of an infrared excess?
Figure~\ref{fig:sed} shows the SED.  While AllWISE \cite{Cutri2014}
yielded a confident W3 detection (9.8$\sigma$), the W4 extraction
yielded only a marginal (1.7$\sigma$) indication of detectable flux.
Similar to other CPVs \cite{Stauffer2017,Bouma2024}, the photometric
uncertainties from WISE W1 and W2 allow at most a $\lesssim$2\% excess
at 3-5\,$\mu$m relative to the stellar photosphere, and a $\lesssim$5\%
excess at 10\,$\mu$m (W3).  A simple approximation is to assume $T_{\rm
dust}$=1500\,K, typical for dust near the star (see \cite{Zhan2019} for
discussion regarding dust sublimation).  We then treat emission from the
dust and the star as Planck functions, requiring $F_{\lambda,\star} > f
\cdot F_{\lambda,\mathrm{dust}}$, for $f$ the scaling factor set by the
WISE photometric uncertainty.  For this condition to hold, we find that
the bolometric flux ratio $F_{\rm dust}/F_{\star}$$<$6$\cdot$10$^{-3}$.
Scaling the associated luminosities ($L = \sigma_{\rm SB} T_{\rm eff}^4
/ (4\pi R^2)$) implies an emitting area for the dust $<$10\% of the
star's surface area.  Converting this constraint to a dust mass requires
an assumption regarding the grain properties; taking a grain density
$\rho$=3\,g\,cm$^{-3}$ typical for silicate grains, and a fixed size
$a$=1\,$\mu$m \cite{Tanimoto2020,Gunther2022,Koen2023} yields an upper
limit of $M_{\rm dust}\lesssim$10$^{17}$\,g.

By way of comparison, the lower limit follows 
%FIXME TODO finish!!

The associated

which implies a 
Since the effective temperatures

luminosity

Under this approximation,

, reporting $W4$=8.4 with no uncertainty.  
The WISE images, available through the IRSA archive, concur.

%WISE_RSR_W3 9.8380 0.0450
%W4 = 8.392 +/- nan
%TODO TODO LEFT OFF HERE
...

@@

\item[Acknowledgments]
  % convos w/
  % M.~Jardine. (line core sum, explaining flux tubes)
  % A Weinberger (dissolving rocks)
  % L Hillenbrand (general?)
  % line profile analysis by B.~Tofflemire.
  % CPS group for reduction support.
  The author thanks X, Y, Z.
  L.G.B. was suported by...
	Acknowledge TESS...
  Acknowledge use of IRSA.
